---
title: "Harry Potter kNN Sorting Hat"
author: "Yahya Gahbiche"
date: "10/1/2020"
output: 
  html_document:
    toc: TRUE
    toc_depth: 3
    toc_float: true
    number_sections: true
---

# Clear environment

```{r}
# Clear environment of variables and functions
rm(list = ls(all = TRUE)) 
```

```{r echo = FALSE, message=FALSE, warning=FALSE, include=FALSE}

# Clear environment of packages
if(is.null(sessionInfo()$otherPkgs) == FALSE)lapply(paste("package:", names(sessionInfo()$otherPkgs), sep=""), detach, character.only = TRUE, unload = TRUE)
```



# Load Libraries:

```{r}
library(FNN)
library(class)
library(ggplot2)
library(lattice)
library(caret)


```


# Load the data

```{r}
hogwarts <- read.csv("super_heroes_hogwarts_v3a.csv", header = TRUE)

# View data, first 10 rows
head(hogwarts,10)

```



## Exploring the data: 

```{r}
# Looking at data types
str(hogwarts)

# Looking at names of different fields
names(hogwarts)

# Number of rows:
nrow(hogwarts)

```

## Removing unnecessary variables

```{r}
hogwarts <- hogwarts[ , -c(1:8, 17, 19:26)]
names(hogwarts) # We will use 6 variables to predict the house

```


## Look at the variables and new order 

```{r}
# transposed
t(t(names(hogwarts))) 

# Exploring dataset: Data types
str(hogwarts)

# number of rows: Still the same
nrow(hogwarts) 

# Viewing all different houses in a table
table(hogwarts$House)
nrow(hogwarts)

```

## Set House as a factor
```{r}
hogwarts$House <- as.factor(hogwarts$House)
```

# Training - Validation split

```{r}
# Set the seed

set.seed(666)
```

+ Randomly sample the rows via their indexes (row number)
We did 70-30 split

```{r}
# Data partitioning

train_index <- sample(1:nrow(hogwarts), 0.7 * nrow(hogwarts)) 
valid_index <- setdiff(1:nrow(hogwarts), train_index)

```


+ Assign the randomly selected indexes to the dataset to create the training and validation sets

```{r}
# Create training and validation datasets

train_df <- hogwarts[train_index, ] 
valid_df <- hogwarts[valid_index, ]

```
It's a good idea to check the datasets before continuing 

```{r}
nrow(train_df)
nrow(valid_df)

head(train_df)
head(valid_df)

str(train_df)
str(valid_df)

```

# Set new Padawan: Me

```{r}

padawan_1 <- data.frame(Manipulative = 2,
                        Resourceful = 8,
                        Dismissive = 2,
                        Intelligent = 7,
                        Trusting = 5,
                        Loyal = 10,
                        Stubborn = 7,
                        Brave = 9)

padawan_1
```

# Normalization

+This is needed if predictors are on a different scale.
+In this case, this is just for illustration.

```{r}
# Assign normalized data to a new dataframe
train_norm <- train_df
valid_norm <- valid_df
```

First, create a normalising algorithm using the 8 variables in the training set.

```{r}
# Just double checking we still have our 8 variables
names(train_df)
```

```{r}
# Normalization algorithm

norm_values <- preProcess(train_df[, -c(9)],
                          method = c("center",
                                     "scale"))

# Normalizing training dataset using train_df

train_norm[, -c(9)] <- predict(norm_values,
                               train_df[, -c(9)])

head(train_norm)
```

+ Then using these normalizing algorithm, predict the normalized values of the validation set.

```{r}
# Normalizing validation dataset

valid_norm[, -c(9)] <- predict(norm_values,
                               valid_df[, -c(9)])


head(valid_norm)
```

+ Finally, predict the normalized values for the new padawan (i.e. the new record)

```{r}
padawan_1_norm <- predict(norm_values, padawan_1)
padawan_1_norm

```
# Optimal k for kNN

First, create a df for accuracy using values from 1 to 19, 1 step at a time.

```{r}
accuracy_df <- data.frame(k = seq(1, 19, 1), accuracy = rep(1, 19))
accuracy_df
```

```{r}
# Still checking on train_norm 

names(train_norm)
head(train_norm)
names(valid_norm)
head(valid_norm)

```
+ Obtain the accuracies using k = i, then generate the output as df. For a large dataset, this can take a while. We can choose k = 8, but to prevent a random tie breaker.
+ This is characteristic of kNN, it’s better to use an odd number k.

```{r}
# Using a loop function to find the best k 

for (i in 1:19) {
  knn_pred <- class::knn(train_norm[, -c(9)],
                         valid_norm[, -c(9)],
                         cl = train_norm[, 9], k = i)
  accuracy_df[i, 2] <- confusionMatrix(knn_pred,
                                       factor(valid_norm[, 9]))$overall[1]
}

accuracy_df
```

+ If you like, plot the accuracies for different k for illustration.

```{r}
# Plotting using ggplot2

ggplot(data = accuracy_df) + aes(x = k, y = accuracy) +
  geom_line() + xlab("Different k") + ylab("Accuracy") +
  ggtitle("Accuracies for different values of k") +
  theme(plot.title = element_text(hjust = 0.5)) +
  annotate("text", x = 10, y = 0.825, label = "k = 7 may be suitable") +
  geom_vline(xintercept = 7, linetype = "dashed", color = "red")
  
```

# kNN for k = 7

```{r}
# Train the kNN model using k = 7

knn_pred_k7 <- class::knn(train = train_norm[, -c(9)],
                          test = valid_norm[, -c(9)],
                          cl = train_df$House,
                          k = 7)
```

```{r}
# Confusion Matrix for the model
confusionMatrix(knn_pred_k7, as.factor(valid_df[, 9]))
```

```{r}
# Associated probability of the prediction

knn_pred_k7_prob <- class::knn(train = train_norm[, -c(9)],
                               test = valid_norm[, -c(9)],
                               cl = train_norm$House,
                               k = 7,
                               prob = TRUE)

knn_pred_k7_prob
```


+ Using the trained model, predict the class (i.e. House) of the validation set.


```{r}
# Predicting outcome using the validation data

knn_pred_k7_valid <- class::knn(train = train_norm[, -c(9)],
                                test = valid_norm[, -c(9)],
                                cl = train_norm[, c(9)],
                                k = 7)


knn_pred_k7_valid
```

```{r}
# Compute the probability.

attr(knn_pred_k7_prob, "prob")
```

# Sorting hat for the new padawan

+ Now predict the class (i.e. House) for the new padawan using the same approach.

+ Compute the probabilities, so we know how probable was each class for this new padawan.

```{r}
# checking names

names(train_norm)
names(padawan_1_norm)


```
```{r}
# Predicting the outcome of the new padawan

padawan_1_pred <- class::knn(train = train_norm[, -c(9)],
                             test = padawan_1_norm,
                             cl = train_norm$House,
                             k =7)

padawan_1_pred

```

+ Now let's do the prob

```{r}
# probability of the padawan outcome
padawan_1_pred_prob <- class::knn(train = train_norm[, -c(9)],
                                  test = padawan_1_norm[, -c(9)],
                                  cl = train_norm$House,
                                  k = 7,
                                  prob = TRUE)

padawan_1_pred_prob

```









